\documentclass{article}
\input{../../preamble.tex}

\usepackage{fancyhdr}
\usepackage{cancel}
\usepackage[margin=1in, headheight=50pt]{geometry}
\pagestyle{fancy}
\lhead{\textbf{Ph22 Set 3}}
\chead{}
\rhead{Aritra Biswas}
\setlength{\headsep}{20pt}

\begin{document}


\section{Restricted three-body problem}

We study an asteroid of mass $m$ in the Jupiter-Sun system (masses $M_1$ and
$M_2$ respectively). Since $m \ll M_1, M_2$, we approximate Jupiter and the
Sun as fixed in the corotating frame, which rotates with angular velocity
$\Omega$:
\begin{align}
    \Omega = \sqrt{G(M_1 + M_2) \over R^3},
    \tsp{where $R_3$ is the Jupiter-Sun separation,}
\end{align}
with respect to an inertial frame. In this frame, Jupiter and the Sun are
fixed at:
\begin{align}
    \vec{r_1} = \left( {M_2 R \over M_1 + M_2}, 0 \right),
    \sp{}
    \vec{r_2} = \left( -{M_1 R \over M_1 + M_2}, 0 \right).
\end{align}
The acceleration of the asteroid at $\vec r = (x, y)$ is given by:
\begin{align}
    \vec{\ddot r} =
    - {GM_1} {\vec r - \vec{r_1} \over |\vec r - \vec{r_1}|^3}
    - {GM_2} {\vec r - \vec{r_2} \over |\vec r - \vec{r_2}|^3}
    + 2\Omega (\dot y, -\dot x)
    + \Omega^2 \vec r,
\end{align}
with gravitational contributions from Jupiter and the Sun, a Coriolis force
contribution, and a centrifugal force contribution. Lagrange's solution
identifies two \emph{stable} Lagrange points $\vec{L_4}$ and $\vec{L_5}$
located at:
\begin{align}
    \vec{L_{4,5}}
    = R \left({M_2 - M_1 \over M_1 + M_2} \cos\alpha_{4,5},
    \sin\alpha_{4,5}\right)
    \sp{where}
    \alpha_{4,5} = \pm {\pi \over 3}.
\end{align}
To observe the stability of these starting points, we will start our asteroid
at nearby points $\vec{\tilde L_{4,5}}$ with $\tilde \alpha_{4, 5}
\equiv \alpha_{4,5} \pm \Delta\alpha$. For small $\Delta\alpha$, we should
see small oscillations around $\vec{L_{4,5}}$, but for large $\Delta\alpha$,
we should see more variation in the asteroid's orbit. We use an RK4
integration routine to evolve the position $\vec r$ of the asteroid over
several periods $T = 2\pi / \Omega$.

Figures \ref{da_0p01.pdf} through \ref{da_1p5.pdf} show asteroid orbits for
$\Delta\alpha = 0.01, 0.05, 0.5, 1.5$. No orbit is shown for
$\Delta\alpha = 0$ because there is none: the asteroid remains stationary
at its starting position. As expected, the asteroid's orbit amplitude
increases dramatically with $\Delta\alpha$.

\plop{da_0p01.pdf}{Small-oscillation orbits of an asteroid in the Jupiter-Sun
    system, placed at starting positions $\vec{\tilde L_{4,5}}$
    with $\Delta\alpha = 0.01$. Yellow star denotes the Sun, and red circle
    denotes Jupiter.}

\plop{da_0p05.pdf}{Small-oscillation orbits of an asteroid in the Jupiter-Sun
    system, placed at starting positions $\vec{\tilde L_{4,5}}$
    with $\Delta\alpha = 0.05$. Yellow star denotes the Sun, and red circle
    denotes Jupiter.}

\plop{da_0p5.pdf}{Larger-oscillation orbits of an asteroid in the Jupiter-Sun
    system, placed at starting positions $\vec{\tilde L_{4,5}}$
    with $\Delta\alpha = 0.5$. Yellow star denotes the Sun, and red circle
    denotes Jupiter.}
 
\plop{da_1p5.pdf}{Larger-oscillation orbits of an asteroid in the Jupiter-Sun
    system, placed at starting positions $\vec{\tilde L_{4,5}}$
    with $\Delta\alpha = 1.5$. Yellow star denotes the Sun, and red circle
    denotes Jupiter.}

\section{General three-body problem with equal masses}

In a general multi-body problem, the acceleration of the $i$-th object is
given by:
\begin{align}
    \vec{\ddot r_i} = -G \sum_{j \not= i} M_j {\vec r_i - \vec r_j
    \over |\vec r_i - \vec r_j|^3}.
\end{align}
We consider a case with three bodies of equal mass. We will use natural
units, where $G = M_{1, 2, 3} = 1$.

\subsection{Lagrange solution}

We want to verify the Lagrange solution, where
the masses move along the vertices of a rotating equilateral triangle of
side length $d$ and velocity $v$. First, we will confirm this solution
analytically.

Let $R \equiv d / \sqrt{3}$, the distance of a point on the equilateral
triangle from the center. This is the radius of the centripetal motion of
each mass. Without loss of generality, let one of the masses begin at
$(x, y) = (0, R)$. By symmetry, the net gravitational force on this mass
is in the $y$-direction (towards the center) and its magnitude is:
\begin{align}
    F = 2 \left(G M^2 \over d^2\right)\cos 30^\circ = 
    {GM^2 \over d^2} \sqrt 3 = {GM^2 \over dR}.
\end{align}
The velocity of centripetal motion is related to the centripetal force by
$F = v^2 / R$, so:
\begin{align}
    v = \sqrt{RF} = \sqrt{GM \over d}.\,\square
\end{align}
We simulate this orbit using our RK4 integrator. Initial conditions are:
\begin{align}
    (x_i, y_i) = (R\sin\theta_i, R\cos\theta_i),
    \sp{} (\dot x_i, \dot y_i) = (v\cos\theta_i, -v\sin\theta_i),
    \sp{with} \{\theta_1, \theta_2, \theta_3\} =
    \left\{0, {2\pi \over 3}, {4\pi \over 3} \right\}.
\end{align}
The result of the simulation is saved in \texttt{lagrange.mp4} in this
directory.

\subsection{Choreographic orbit}

Using the same RK4 routine, we use the given conditions for a
choreographic orbit. The simulation result is saved in
\texttt{choreographic.mp4} in this directory.



\end{document}
